---
title: "DSI"
author: "whr"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1.load package

```{r}
library(haven) # read .dta file
library(dplyr)
library(ggplot2)
library(stringr)
library(patchwork)
library(broom)
```
# 2.read data

```{r}
data_mtlv <- read_dta("data/mat_leave_country_year.dta")
data_other_vars <- read.csv('data/final_dataset1.csv')
```


## 2.1. tidy country name

```{r}
country_mapping <- data.frame(
  code = c("AM", "BD", "BF", "BJ", "CG", "CM", "CO", "DR", "ET", "GA", "GH", "GN", "HT", "JO", "KE", "KH", "LS", "MD", "ML", "MZ", "NG", "NI", "NM", "NP", "PE", "RW", "SN", "TZ", "UG", "ZM", "ZW"),
  full_name = c("Armenia", "Bangladesh", "Burkina Faso", "Benin", "Congo", "Cameroon", "Colombia", "Dominican Republic", 
                "Ethiopia", "Gabon", "Ghana", "Guinea", "Haiti", "Jordan", "Kenya", "Cambodia", "Lesotho", "Moldova", 
                "Mali", "Mozambique", "Nigeria", "Nicaragua", "Namibia", "Nepal", "Peru", "Rwanda", "Senegal", 
                "Tanzania", "Uganda", "Zambia", "Zimbabwe")
)

data_other_vars <- data_other_vars %>%
  left_join(country_mapping, by = c("country" = "code")) %>%
  mutate(country = full_name) %>% 
  select(-full_name)
```


## 2.2.common country

```{r}
common_countries <- intersect(unique(data_other_vars$country), unique(data_mtlv$country))
common_countries
```



## 2.3.treated country


```{r}
treated_data <- data_other_vars[data_other_vars$status == "Treated", ]
treated_countries <- unique(treated_data$country)
print(treated_countries)
```


## 2.4.missing country
```{r}
missing_country1 <- setdiff(unique(data_other_vars$country), unique(data_mtlv$country))
missing_country2 <- setdiff(unique(data_mtlv$country), unique(data_other_vars$country))

print(missing_country1)
print(missing_country2)
```

## 2.5.data only common country

```{r}
data_other_vars_common <- data_other_vars %>%
  filter(country %in% common_countries)

data_mtlv_common <- data_mtlv %>%
  filter(country %in% common_countries)
```


# 3.mtlv

## 3.1.mtlv mean by country

```{r}
data_mtlv_mean <- data_mtlv %>%
  group_by(country) %>%
  summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE)))
```

## 3.2.mtlv mean only common country

```{r}
data_mtlv_mean_common <- data_mtlv_mean %>%
  filter(country %in% common_countries)
```

## 3.3.plot mtlv mean

```{r}
data_mtlv_long <- data_mtlv_mean_common %>%
  tidyr::pivot_longer(
    cols = starts_with("mtlv_pdr_"),  # Select the column whose name begins with “mtlv_pdr_”
    names_to = "year",                # New column: Year column
    values_to = "value"               # New column: column for indicator values
  ) %>%
  mutate(
    year = as.numeric(sub("mtlv_pdr_", "", year))  # Extract year numbers and convert to numeric types
  )
```


```{r}
ggplot(data_mtlv_long, aes(x = year, y = value, color = country, group = country)) +
  geom_line() +       # Adding Curves
  geom_point() +      # Adding Data Points
  labs(
    title = "Country-wise Changes Over Years",
    x = "Year",
    y = "Value",
    color = "Country"
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",         # Setting the legend position
    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotating year labels
  )
```



# 4.other variable data

## 4.1.variable selection

```{r}
analysis_variable <- data_other_vars_common %>% 
  select(hforage_mean, # HAZ, cts var
         stunted_mean, # Degree of stunting
         weight_mean,  # cluster weight mean
         rural_mean,   # if rural (0 or 1)
         mat_lit_mean, # clusterMaternal literacy rate
         mat_height_mean, # cluster Average height of mothers
         matwork_mean, # clusterProportion of working mothers
         country,
         status,
         survey_year
         )
```


## 4.2.mean by country

```{r}
analysis_variable_mean <- analysis_variable %>%
  group_by(country) %>%
  summarise(across(where(is.numeric), \(x) mean(x, na.rm = TRUE)))
```


## 4.3.into blocks

```{r}
split_into_blocks <- function(df, block_size = 500) {
  df <- df %>%
    mutate(block = ceiling(row_number() / block_size))
  return(df)
}

# Grouping and slicing by country block
data_with_blocks <- analysis_variable %>%
  group_by(country,
           survey_year) %>%
  group_modify(~ split_into_blocks(.x, block_size = 500)) %>%
  ungroup()
```

## 4.4.mean by block

```{r}
block_means <- data_with_blocks %>%
  group_by(country, 
           survey_year,
           block) %>%
  summarise(across(where(is.numeric), mean, na.rm = TRUE)) %>%
  ungroup()
```


# 5.modeling

## 5.1.PCA



```{r}
# Remove rows containing NA
block_means_clean <- block_means %>%
  na.omit()

# Remove non-numeric columns and keep only the numeric part for dimensionality reduction
block_means_numeric <- block_means_clean %>%
  select(-country, -block)
```


```{r}
pca_result <- prcomp(block_means_numeric, scale. = TRUE)

# Extracting the first two principal components
pca_2d <- as.data.frame(pca_result$x[, 1:2])
colnames(pca_2d) <- c("PC1", "PC2")  # 重命名列

# Combining downscaling results with original country and block information
pca_2d <- cbind(block_means_clean[, c("country","survey_year", "block")], pca_2d)
```


## 5.2.scree plot

```{r}
explained_variance <- pca_result$sdev^2 / sum(pca_result$sdev^2)

```


```{r}
scree_data <- data.frame(
  Principal_Component = seq_along(explained_variance),
  Explained_Variance = explained_variance
)
```


```{r}
ggplot(scree_data, aes(x = Principal_Component, y = Explained_Variance)) +
  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.7) +
  geom_line(color = "red", linewidth = 1) +
  geom_point(color = "red", size = 2) +
  labs(title = "Scree Plot",
       x = "Principal Component",
       y = "Proportion of Variance Explained") +
  theme_minimal()
```


## 5.3.scatter plot by country

```{r}
ggplot(pca_2d, aes(x = PC1, y = PC2, color = country)) +
  geom_point(size = 3, alpha = 0.8) +  # Scatter size and transparency
  labs(title = "PCA: Block-Level Averages (2D Projection)",
       x = "Principal Component 1",
       y = "Principal Component 2") +
  theme_minimal() +  # 
  theme(legend.title = element_text(size = 12),  # Legend title font
        legend.text = element_text(size = 10)) 
```

## 5.4.scatter plot by status

```{r}
status_info <- data_with_blocks %>%
  select(country, block, status) %>%
  distinct()

# Merge status information into PCA data
pca_2d_with_status <- pca_2d %>%
  left_join(status_info, by = c("country", "block"))

```


```{r}
ggplot(pca_2d_with_status, aes(x = PC1, y = PC2, color = status)) +
  geom_point(size = 3, alpha = 0.8) +
  labs(title = "PCA: Block-Level Averages",
       x = "Principal Component 1",
       y = "Principal Component 2",
       color = "Group") +
  theme_minimal()
```



```{r}
country_year_count <- data_other_vars_common %>%
  group_by(country) %>%
  summarise(num_survey_years = n_distinct(survey_year)) %>%
  arrange(desc(num_survey_years))  # Sort by number in descending order


print(country_year_count)

country_years <- data_other_vars_common %>%
  group_by(country) %>%
  summarise(survey_years = paste(sort(unique(survey_year)), collapse = ", ")) %>%
  arrange(country)


print(country_years)
```


```{r}
common_countries
treated_countries
```

## 5.5. 4 groups
```{r}
pca_2d_4groups <- pca_2d_with_status %>%
  mutate(group = case_when(
    status == "Control" & survey_year < 2004 ~ "Control: Before 2004",
    status == "Control" & survey_year >= 2004 ~ "Control: After 2004",
    status == "Treated" & survey_year < 2004 ~ "Treated: Before 2004",
    status == "Treated" & survey_year >= 2004 ~ "Treated: After 2004"
  ))

```


```{r}

group_colors <- c(
  "Control: Before 2004" = "salmon",
  "Control: After 2004" = "red",
  "Treated: Before 2004" = "#a2d2ff",
  "Treated: After 2004" = "blue"
)

ggplot(pca_2d_4groups, aes(x = PC1, y = PC2, color = group)) +
  geom_point(size = 2, alpha = 0.8) +
  scale_color_manual(values = group_colors) + 
  labs(title = "Cluster 4: maternity leave removed",
       x = "Principal Component 1",
       y = "Principal Component 2",
       color = "Group") +  # Legend title
  theme_minimal() +
  theme(
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10)
  )
```
```{r}
p <- ggplot(pca_2d_4groups, aes(x = PC1, y = PC2, color = group)) +
  geom_point(size = 2, alpha = 0.8) +
  scale_color_manual(values = group_colors) +
  labs(title = "Cluster 4: maternity leave removed",
       x = "Principal Component 1",
       y = "Principal Component 2",
       color = "Group") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10)
  )

ggsave(
  filename = "DSI Cluster 4.png",
  plot = p,
  width = 8,
  height = 6,
  dpi = 300
)
```




# 6.merge data

## 6.1. merge

```{r}
colnames(data_mtlv_mean_common) <- str_remove(colnames(data_mtlv_mean_common), "mtlv_pdr_")
```

```{r}
data_mtlv_long <- data_mtlv_mean_common %>%
  tidyr::pivot_longer(
    cols = -country,  # Convert all columns except the “country” column to long form
    names_to = "year",  # New column name “year”
    values_to = "maternity_leave"  # New listing for length of maternity leave
  ) %>%
  mutate(year = as.numeric(year))
```


```{r}
merged_data <- data_other_vars_common %>%
  left_join(
    data_mtlv_long,
    by = c("country" = "country", "survey_year" = "year")
  )
```


## 6.2. select

### 6.2.1.all variable
```{r}
merged_data_selected <- merged_data %>% 
  select(hforage_mean, # HAZ, cts var
         stunted_mean,
         weight_mean,  
         rural_mean,  
         mat_lit_mean, 
         mat_height_mean, 
         matwork_mean,
         country,
         status,
         survey_year,
         maternity_leave
         )
```





### 6.2.2.into block

```{r}
split_into_blocks <- function(df, block_size = 500) {
  df <- df %>%
    mutate(block = ceiling(row_number() / block_size))
  return(df)
}

# Grouping and slicing by country block
merged_selected_blocks <- merged_data_selected %>%
  group_by(country,
           survey_year) %>%
  group_modify(~ split_into_blocks(.x, block_size = 500)) %>%
  ungroup() %>% 
  group_by(country, 
           survey_year,
           block) %>%
  summarise(across(where(is.numeric), mean, na.rm = TRUE)) %>%
  ungroup()
```

```{r}
# define matching library
status_vector <- setNames(merged_data_selected$status, merged_data_selected$country)

# use match() to add `status` column to merged_selected_blocks
merged_selected_blocks$status <- status_vector[merged_selected_blocks$country]
```


### 6.2.3.clean NA

```{r}
merged_selected_blocks <- merged_selected_blocks %>%
  na.omit()
```


### 6.2.4. 4groups

```{r}
merged_selected_blocks <- merged_selected_blocks %>%
  mutate(group = case_when(
    status == "Control" & survey_year < 2004 ~ "Control: Before 2004",
    status == "Control" & survey_year >= 2004 ~ "Control: After 2004",
    status == "Treated" & survey_year < 2004 ~ "Treated: Before 2004",
    status == "Treated" & survey_year >= 2004 ~ "Treated: After 2004"
  ))

```



# 7. DID


## 7.1.did term
```{r}
merged_selected_blocks_did <- merged_selected_blocks %>%
  mutate(
    Post = ifelse(survey_year >= 2004, 1, 0),  # 1 for 2004 and onwards, 0 before that
    Treatment = ifelse(status == "Treated", 1, 0),  # 1 for the Treated group and 0 for the Control group.
    DID = Post * Treatment  # 交互项
  )
```

## 7.2.dependent variable

```{r}
merged_dep_var <- merged_selected_blocks_did %>% 
  select(hforage_mean,
         stunted_mean, 
         weight_mean  
         )
```


### 7.2.1.PCA

```{r}
pca_result <- prcomp(merged_dep_var, scale. = TRUE)

# Extracting the first two principal components
pca_dep_var <- as.data.frame(pca_result$x[, 1])
colnames(pca_dep_var) <- c("PC1")  # rename columns

# Combining results with original country and block information
#pca_2d <- cbind(merged_selected_blocks_did[, c("country","survey_year", "status")], pca_dep_var)
```

### 7.2.2.scree plot


```{r}
explained_variance <- pca_result$sdev^2 / sum(pca_result$sdev^2)

```


```{r}
scree_data <- data.frame(
  Principal_Component = seq_along(explained_variance),
  Explained_Variance = explained_variance
)
```


```{r}
ggplot(scree_data, aes(x = Principal_Component, y = Explained_Variance)) +
  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.7) +
  geom_line(color = "red", linewidth = 1) +
  geom_point(color = "red", size = 2) +
  labs(title = "Scree Plot",
       x = "Principal Component",
       y = "Proportion of Variance Explained") +
  theme_minimal()
```

## 7.2.independent variable

```{r}
merged_indep_var <- merged_selected_blocks_did %>% 
  select(rural_mean,  
         mat_lit_mean,
         mat_height_mean, 
         matwork_mean, 
         maternity_leave,
         Post,
         Treatment,
         DID
         )
```


## 7.3. merge data

```{r}
de_inde_pca <- cbind(pca_dep_var, merged_indep_var)
```

## 7.4.DID model

### 7.4.1.M1: simple

```{r}
did_model_1 <- lm(PC1 ~ Post * Treatment, data = de_inde_pca)

summary(did_model_1)
```

### 7.4.2.M2: add mother factor

```{r}
did_model_2 <- lm(PC1 ~ Post * Treatment + mat_lit_mean + mat_height_mean + matwork_mean, data = de_inde_pca)

summary(did_model_2)
```

### 7.4.3.M3: add mother factor and rural


```{r}
did_model_3 <- lm(PC1 ~ Post * Treatment + mat_lit_mean + mat_height_mean +matwork_mean + rural_mean, data = de_inde_pca)

summary(did_model_3)
```


### 7.4.4.Table


```{r}
extract_model_info <- function(model, model_name) {
  # Extraction coefficients and confidence intervals
  coef_info <- tidy(model, conf.int = TRUE) %>%
    filter(term == "Post:Treatment") %>%
    select(estimate, p.value, conf.low, conf.high)
  
  # Extract R^2
  r_squared <- summary(model)$r.squared
  
  # Return results
  tibble(
    Model = model_name,
    Estimate = coef_info$estimate,
    `P-value` = coef_info$p.value,
    `95% CI` = paste0("(", round(coef_info$conf.low, 3), ", ", round(coef_info$conf.high, 3), ")"),
    `R^2` = r_squared
  )
}

```


```{r}
# Extracting information for each model
model1_info <- extract_model_info(did_model_1, "Model 1")
model2_info <- extract_model_info(did_model_2, "Model 2")
model3_info <- extract_model_info(did_model_3, "Model 3")

# Merge results into a table
result_table <- bind_rows(model1_info, model2_info, model3_info)

print(result_table)
```

```{r}
plot <- ggplot(result_table, aes(x = Model, y = Estimate)) +
  geom_point(size = 3) +  # Plotting points indicates the value of the coefficient
  geom_errorbar(aes(ymin = as.numeric(sub("\\(", "", gsub(",.*", "", `95% CI`))), 
                    ymax = as.numeric(sub(".*,", "", gsub("\\)", "", `95% CI`)))), 
                width = 0.2) +  # confidence interval 
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +  # zero base line 
  geom_text(aes(label = round(Estimate, 3)), vjust = -1, size = 4, color = "black") +  # Labeling coefficient values
  labs(
    title = "Effect of Post × Treatment Across Models",
    x = "Model",
    y = "Estimate (Post × Treatment)"
  ) +
  theme_minimal()

plot <- plot +
  geom_text(aes(label = paste0("p = ", round(`P-value`, 3), "\nR² = ", round(`R^2`, 3))), 
            vjust = -1, size = 4, color = "blue")
print(plot)


ggsave("DID_model_results.png", plot = plot, width = 8, height = 5)

```



# 8. cluster

## 8.1.dependent variables

```{r}
pca_result <- prcomp(merged_dep_var, scale. = TRUE)

pca_dep_var <- as.data.frame(pca_result$x[, 1:2])
colnames(pca_dep_var) <- c("PC1","PC2")  

de_var_cluster <- cbind(merged_selected_blocks_did[, c("country","survey_year", "status","group")], pca_dep_var)
```


```{r}
group_colors <- c(
  "Control: Before 2004" = "salmon",
  "Control: After 2004" = "red",
  "Treated: Before 2004" = "#a2d2ff",
  "Treated: After 2004" = "blue"
)

ggplot(de_var_cluster, aes(x = PC1, y = PC2, color = group)) +
  geom_point(size = 2, alpha = 0.8) +
  scale_color_manual(values = group_colors) + 
  labs(title = "PCA Scatter Plot: Grouped by Survey Year and Status",
       x = "Principal Component 1",
       y = "Principal Component 2",
       color = "Group") +  
  theme_minimal() +
  theme(
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10)
  )
```


```{r}
p <- ggplot(de_var_cluster, aes(x = PC1, y = PC2, color = group)) +
  geom_point(size = 2, alpha = 0.8) +
  scale_color_manual(values = group_colors) + 
  labs(title = "Cluster 1: dependent variables",
       x = "Principal Component 1",
       y = "Principal Component 2",
       color = "Group") +  
  theme_minimal() +
  theme(
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10)
  )


ggsave(
  filename = "DSI Cluster 1.png",
  plot = p,
  width = 8,
  height = 6,
  dpi = 300
)
```



## 8.2.independent variables

```{r}
inde_var_cluster <- merged_selected_blocks_did %>% 
  select(rural_mean,   
         mat_lit_mean, 
         mat_height_mean, 
         matwork_mean
         )
```


```{r}
pca_result <- prcomp(inde_var_cluster, scale. = TRUE)

pca_indep_var <- as.data.frame(pca_result$x[, 1:2])
colnames(pca_indep_var) <- c("PC1","PC2")  

inde_var_cluster <- cbind(merged_selected_blocks_did[, c("country","survey_year", "status","group")], pca_indep_var)
```

```{r}
group_colors <- c(
  "Control: Before 2004" = "salmon",
  "Control: After 2004" = "red",
  "Treated: Before 2004" = "#a2d2ff",
  "Treated: After 2004" = "blue"
)

ggplot(inde_var_cluster, aes(x = PC1, y = PC2, color = group)) +
  geom_point(size = 2, alpha = 0.8) +
  scale_color_manual(values = group_colors) +  # Using custom colors
  labs(title = "PCA Scatter Plot: Grouped by Survey Year and Status",
       x = "Principal Component 1",
       y = "Principal Component 2",
       color = "Group") +  # Legend title
  theme_minimal() +
  theme(
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10)
  )
```

```{r}
p <- 
ggplot(inde_var_cluster, aes(x = PC1, y = PC2, color = group)) +
  geom_point(size = 2, alpha = 0.8) +
  scale_color_manual(values = group_colors) +  # Using custom colors
  labs(title = "Cluster 2: independent variables",
       x = "Principal Component 1",
       y = "Principal Component 2",
       color = "Group") +  # Legend title
  theme_minimal() +
  theme(
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10)
  )

ggsave(
  filename = "DSI Cluster 2.png",
  plot = p,
  width = 8,
  height = 6,
  dpi = 300
)
```




## 8.3.all variable

```{r}
all_var_cluster <- merged_selected_blocks_did %>% 
  select(hforage_mean, 
         stunted_mean,
         weight_mean,  
         rural_mean,  
         mat_lit_mean, 
         mat_height_mean, 
         matwork_mean 
         )
```


```{r}
pca_result <- prcomp(all_var_cluster, scale. = TRUE)

pca_indep_var <- as.data.frame(pca_result$x[, 1:2])
colnames(pca_indep_var) <- c("PC1","PC2")  

inde_var_cluster <- cbind(merged_selected_blocks_did[, c("country","survey_year", "status","group")], pca_indep_var)
```

```{r}
group_colors <- c(
  "Control: Before 2004" = "salmon",
  "Control: After 2004" = "red",
  "Treated: Before 2004" = "#a2d2ff",
  "Treated: After 2004" = "blue"
)

ggplot(inde_var_cluster, aes(x = PC1, y = PC2, color = group)) +
  geom_point(size = 2, alpha = 0.8) +
  scale_color_manual(values = group_colors) +  
  labs(title = "PCA Scatter Plot: Grouped by Survey Year and Status",
       x = "Principal Component 1",
       y = "Principal Component 2",
       color = "Group") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10)
  )
```

```{r}
p <- ggplot(inde_var_cluster, aes(x = PC1, y = PC2, color = group)) +
  geom_point(size = 2, alpha = 0.8) +
  scale_color_manual(values = group_colors) + 
  labs(title = "Cluster 3: all variables",
       x = "Principal Component 1",
       y = "Principal Component 2",
       color = "Group") +
  theme_minimal() +
  theme(
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10)
  )

ggsave(
  filename = "DSI Cluster 3.png",
  plot = p,
  width = 8,
  height = 6,
  dpi = 300
)
```

















